---
title: "STAT672_ Time Series Project"
output:
  html_document:
    df_print: paged
---

# 0 Load Libraries
```{r}
library(usethis)
library(devtools)
library(ggplot2)
library(ggfortify)
library(dplyr)
library(magrittr)
library(lubridate)
library(forecast)
library(tidyverse)
library(mt)
library(lmtest)
library(outliers)
library(EnvStats)

```


```{r}
library(tseries)
```


# 1 Load Data for station 1 
```{r}
getwd()
```

```{r}
oiltemp_1 <- read.csv("./oil station 1 hourly average for each day.csv",header=TRUE)
head(oiltemp_1, 10)
```

```{r}
attach(oiltemp_1)
```
```{r}
#summary(oiltemp_1)
```
```{r}
colnames(oiltemp_1)
```
```{r}
#Feature_X<-as.matrix(cbind(oiltemp_1$HUFL,oiltemp_1$HULL, oiltemp_1$MUFL, oiltemp_1$MULL, oiltemp_1$LUFL, oiltemp_1$LULL, oiltemp_1$OT))
#par(pairs(Feature_X))
```
```{r}
#pca <- prcomp(Feature_X, scale=TRUE)
#summary(pca)
```
```{r}
colnames(oiltemp_1)
```

```{r}
save.plot <- ggplot(data=oiltemp_1, aes(x=LULL, y=OT))
save.plot + geom_point(size=3)

save.plot <- ggplot(data=oiltemp_1, aes(x=LUFL, y=OT))
save.plot + geom_point(size=3)

save.plot <- ggplot(data=oiltemp_1, aes(x=MULL, y=OT))
save.plot + geom_point(size=3)

save.plot <- ggplot(data=oiltemp_1, aes(x=MUFL, y=OT))
save.plot + geom_point(size=3)

save.plot <- ggplot(data=oiltemp_1, aes(x=HULL, y=OT))
save.plot + geom_point(size=3)

save.plot <- ggplot(data=oiltemp_1, aes(x=HUFL, y=OT))
save.plot + geom_point(size=3)
```

## Predictive Variable
```{r}
#qqPlot(OT)

ggplot(oiltemp_1, aes(x=TimeID)) + 
  geom_line(aes(y=OT), color="darkblue")

#colnames(oiltemp_1)
```

```{r}
acf(OT, lag.max=30)
```

```{r}
pacf(OT, lag.max=30)
```

```{r}
adf.test(OT)
```


## Feature

```{r}
ggplot(oiltemp_1, aes(x=TimeID)) + 
  geom_line(aes(y=HUFL), color="darkblue")

ggplot(oiltemp_1, aes(x=TimeID)) + 
  geom_line(aes(y=HULL), color="darkblue")

ggplot(oiltemp_1, aes(x=TimeID)) + 
  geom_line(aes(y=MUFL), color="darkblue")

ggplot(oiltemp_1, aes(x=TimeID)) + 
  geom_line(aes(y=MULL), color="darkblue")

ggplot(oiltemp_1, aes(x=TimeID)) + 
  geom_line(aes(y=LUFL), color="darkblue")

ggplot(oiltemp_1, aes(x=TimeID)) + 
  geom_line(aes(y=LULL), color="darkblue")
```


```{r}
adf.test(HUFL)
adf.test(HULL)
adf.test(MUFL)
adf.test(MULL)
adf.test(LUFL)
adf.test(LULL)
```



```{r}
acf(HUFL, lag.max=30)
acf(HULL, lag.max=30)
acf(MUFL, lag.max=30)
acf(MULL, lag.max=30)
acf(LUFL, lag.max=30)
acf(LULL, lag.max=30)
```





# 2 Classical Regression 

fit1 <- lm(xt1 ~ time(xt1))
summary(fit1)
anova(fit1)
plot(xt1, type="o", ylab="Oil Temperature 1")
abline(fit1)
plot(as.vector(time(xt1)), fit1$res, xlab="Days")
abline(h=0)


oiltemp_1$HUFL + oiltemp_1$HULL + oiltemp_1$MUFL + oiltemp_1$MULL + oiltemp_1$LUFL + oiltemp_1$LULL
fit2 <- lm(xt1 ~ oiltemp_1$HUFL + oiltemp_1$HULL + oiltemp_1$MUFL + oiltemp_1$MULL + oiltemp_1$LUFL + oiltemp_1$LULL + time(xt1))
summary(fit2)
anova(fit2)
plot(xt1, type="o", ylab="Oil Temperature 1")
abline(fit2)

fit3 <- lm(xt1 ~ oiltemp_1$MUFL + oiltemp_1$LUFL + time(xt1))
summary(fit3)
plot(xt1, type="o", ylab="Oil Temperature 1")
abline(fit3)



```{r}
colnames(oiltemp_1)
```


## Model fit linear regression 
### 1) model1
```{r}
trend <- time(OT)
hufl <- HUFL - mean(HUFL)
hull <- HULL - mean(HULL)
mufl <- MUFL - mean(MUFL)
mull <- MULL - mean(MULL)
lufl <- LUFL - mean(LUFL)
lull <- LULL - mean(LULL)
lufl2 <- lufl^2
lull2 <- lull^2
fit.model1 <- lm(OT ~ trend + hufl + hull + mufl + mull + (lufl2) + (lull2))
```

```{r}
summary(fit.model1)
```

```{r}
anova(fit.model1)
```

```{r}
plot(OT, type="line")
lines(fit.model1$fitted.values, lty=2, col="blue")
```

```{r}
num <- length(OT)
# AIC
AIC(fit.model1)/num - log(2*pi)
# BIC
AIC(fit.model1, k=log(num))/num - log(2*pi)
# AICc
log(sum(resid(fit.model1)^2)/num) + (num +5)/(num -5 -2)
```

check residual assumption: iid, normality
```{r}
plot(as.vector(time(OT)), fit.model1$residual, xlab="Days")
abline(h=0)
```
residual assumption: independence \ 

H0: The data are independently distributed 
Ha: The data are not independently distributed; they exhibit serial correlation.
```{r}
# Box-Pierce test and Ljung-Box test
Box.test(fit.model1$residuals, lag=40, fitdf=2)
Box.test(fit.model1$residuals, lag=40, fitdf=1)
Box.test(fit.model1$residuals, lag=40, fitdf=0)
```

```{r}
acf(fit.model1$residuals, lag.max=40)
rho.hat <- acf(fit.model1$residuals, lag.max=40, plot=F)$acf[-1]
n <- length(fit.model1$residuals)
h <- length(rho.hat)
q <- n*(n+2)*sum(rho.hat^2/(n-(1:h)))
q
pchisq(q, df=h, lower=F)

```

Check normality
```{r}
qqnorm(fit.model1$residuals)
hist(fit.model1$residuals)
```



### 2) model2
```{r}
trend <- time(OT)
hufl <- HUFL - mean(HUFL)
hull <- HULL - mean(HULL)
mufl <- MUFL - mean(MUFL)
mull <- MULL - mean(MULL)
lufl <- LUFL - mean(LUFL)
lull <- LULL - mean(LULL)

lufl2 <- lufl^2
lull2 <- lull^2

fit.model2 <- lm(OT ~ trend + hufl + hull + mufl + mull + (lufl +lufl2) + (lull +lull2))
```

```{r}
summary(fit.model2)
```

```{r}
anova(fit.model2)
```


```{r}
plot(OT, type="line")
lines(fit.model2$fitted.values, lty=2, col="blue")
```


```{r}
num <- length(OT)
# AIC
AIC(fit.model2)/num - log(2*pi)
# BIC
AIC(fit.model2, k=log(num))/num - log(2*pi)
# AICc
log(sum(resid(fit.model2)^2)/num) + (num +5)/(num -5 -2)
```



```{r}
# Box-Pierce test and Ljung-Box test
Box.test(fit.model2$residuals, lag=40, fitdf=2)
Box.test(fit.model2$residuals, lag=40, fitdf=1)
Box.test(fit.model2$residuals, lag=40, fitdf=0)
```

```{r}
qqnorm(fit.model2$residuals)
hist(fit.model2$residuals)
```




### 3) model3
```{r}
trend <- time(OT)
hufl <- HUFL - mean(HUFL)
hull <- HULL - mean(HULL)
mufl <- MUFL - mean(MUFL)
mull <- MULL - mean(MULL)
lufl <- LUFL - mean(LUFL)
lull <- LULL - mean(LULL)

lufl2 <- lufl^2
lull2 <- lull^2

hufl2 <- hufl^2
hull2 <- hull^2

mufl2 <- mufl^2
mull2 <- mull^2

fit.model3 <- lm(OT ~ trend + (hufl+hufl2) + (hull+hull2) + (mufl+mufl2) + (mull+mull2) + (lufl+lufl2) + (lull+lull2))
```

```{r}
summary(fit.model3)
```
```{r}
anova(fit.model3)
```


```{r}
num <- length(OT)
# AIC
AIC(fit.model3)/num - log(2*pi)
# BIC
AIC(fit.model3, k=log(num))/num - log(2*pi)
# AICc
log(sum(resid(fit.model3)^2)/num) + (num +5)/(num -5 -2)
```




```{r}
plot(OT)
lines(fit.model3$fitted.values, lty=2, col="blue")
```






# 3 Data Process

## 3.1 Stationary Checking

Log Transformation 

```{r}
ggplot(oiltemp_1, aes(x=TimeID)) + 
  geom_line(aes(y=log(OT)), color="darkblue")
```

```{r}
plot(log(OT), type="line")
```


```{r}
log.OT <- na.remove(log(OT))
```

```{r}
adf_test <- adf.test(log.OT)
adf_test
```





```{r}
adf_test <- adf.test(OT)
adf_test
```





## 3.2 Differencing to remove trend and seasonality

The first step in the analysis of any time series is to plot the data. \
1) if there are any apparent discontinuities in the series, such as a sudden change of level, it may be advisable to analyze the series by first breaking it into homogeneous segments. \
2) if there are outlying observations, they should be studied carefully to check whether there is any justification for discarding them. \


```{r}
plot(OT, type="line", col="blue")
plot(diff(OT), type="line", col="green")
plot(diff(log(OT)), type="line")
```


```{r}
acf(OT)
acf(diff(OT))
acf(diff(log.OT))
```

```{r}
mean(diff(OT))
```

## 3.3 predictive variable after trend removed 

```{r}
Y <- diff(OT)
```



```{r}
t <- time(Y)
t.sq <- t^2
fit.model4 <- lm(Y ~ t)
```

```{r}
summary(fit.model4)
```


```{r}
plot(Y, type="line", lwd=1, col="blue")
```

```{r}
acf(Y, lag.max = 40)
```

```{r}
pacf(Y, lag.max = 40)
```



# 4 Building Models: AR() MA() ARIMA()

Fit the model, \
Forecating, \
Check Residuals, \



## 4.1 AR(p=8) Note12 Dow Jones Utilities

by pacf, cuts off after lag(8)

```{r}
#adfTest(Y, lag=4, type="nc")
adf.test(Y)
```


```{r}
model.fit1 <- arima(pred, order=c(8, 0, 0))
model.fit1
```

```{r}
num <- length(OT)
# AIC
AIC(model.fit1)/num - log(2*pi)
# BIC
AIC(model.fit1, k=log(num))/num - log(2*pi)
# AICc
log(sum(resid(model.fit1)^2)/num) + (num +5)/(num -5 -2)
```



```{r}
model1 <- predict(model.fit1, n.ahead=10)
model1

n <- length(Y)
L <- min(Y)
U <- max(Y)
plot(Y, xlim=c(1, n+3), ylim=c(L,U), type="line")
abline(h=0)

lines(model1$pred, col="red", lwd=2)
lines((model1$pred + 1.96*model1$se), col="red", lty=2, lwd=2)
lines((model1$pred - 1.96*model1$se), col="red", lty=2, lwd=2)
```


check residual assumption: iid, normality
```{r}
plot(as.vector(time(Y)), model.fit1$residual, xlab="Days")
abline(h=0)
```
residual assumption: independence \ 

H0: The data are independently distributed 
Ha: The data are not independently distributed; they exhibit serial correlation.
```{r}
# Box-Pierce test and Ljung-Box test
Box.test(model.fit1$residuals, lag=40, fitdf=2)
Box.test(model.fit1$residuals, lag=40, fitdf=1)
Box.test(model.fit1$residuals, lag=40, fitdf=0)
```

```{r}
acf(model.fit1$residuals, lag.max=40)
rho.hat <- acf(model.fit1$residuals, lag.max=40, plot=F)$acf[-1]
n <- length(model.fit1$residuals)
h <- length(rho.hat)
q <- n*(n+2)*sum(rho.hat^2/(n-(1:h)))
q
pchisq(q, df=h, lower=F)

```

Check normality
```{r}
qqnorm(fit.model1$residuals)
hist(fit.model1$residuals)
```







## 4.2 MA(q=2) Note12 Overshorts

by acf, cuts off after lag(2) 

```{r}
model.fit2 <- arima(Y, order=c(0, 0, 2))
model.fit2
```




```{r}
num <- length(OT)
# AIC
AIC(model.fit2)/num - log(2*pi)
# BIC
AIC(model.fit2, k=log(num))/num - log(2*pi)
# AICc
log(sum(resid(model.fit2)^2)/num) + (num +5)/(num -5 -2)
```




```{r}
model2 <- predict(model.fit2, n.ahead=10)
model2
```

```{r}
n <- length(Y)
L <- min(Y)
U <- max(Y)
plot(Y, xlim=c(1, n+3), ylim=c(L,U), type="line")
abline(h=0)

lines(model2$pred, col="red", lwd=2)
lines((model2$pred + 1.96*model2$se), col="red", lty=2, lwd=2)
lines((model2$pred - 1.96*model2$se), col="red", lty=2, lwd=2)
```



check residual assumption: iid, normality
```{r}
plot(as.vector(time(Y)), model.fit2$residual, xlab="Days")
abline(h=0)
```

residual assumption: independence \ 

H0: The data are independently distributed 
Ha: The data are not independently distributed; they exhibit serial correlation.
```{r}
# Box-Pierce test and Ljung-Box test
Box.test(model.fit2$residuals, lag=40, fitdf=2)
Box.test(model.fit2$residuals, lag=40, fitdf=1)
Box.test(model.fit2$residuals, lag=40, fitdf=0)
```

```{r}
acf(model.fit2$residuals, lag.max=40)
rho.hat <- acf(model.fit2$residuals, lag.max=40, plot=F)$acf[-1]
n <- length(model.fit2$residuals)
h <- length(rho.hat)
q <- n*(n+2)*sum(rho.hat^2/(n-(1:h)))
q
pchisq(q, df=h, lower=F)

```

Check normality
```{r}
qqnorm(fit.model2$residuals)
hist(fit.model2$residuals)
```






## 4.3 Order Selection Note13

sample ACF and PACF plots are of little use when both p>0 and q>0 in ARMA model. \
we need to find a systematic way to choose p and q. Assuming white noise normally distributed, we choose the model which has the smallest AICc value. \ 


```{r}
library(forecast)
#auto.arima(Y)
auto <- auto.arima(Y, d=0, ic="aicc", trace=T)
auto

```




```{r}
num <- length(OT)
# AIC
AIC(auto)/num - log(2*pi)
# BIC
AIC(auto, k=log(num))/num - log(2*pi)
# AICc
log(sum(resid(auto)^2)/num) + (num +5)/(num -5 -2)
```




```{r}
model2 <- predict(auto, n.ahead=10)
model2
```

```{r}
n <- length(Y)
L <- min(Y)
U <- max(Y)
plot(Y, xlim=c(1, n+3), ylim=c(L,U), type="line")
abline(h=0)

lines(model2$pred, col="red", lwd=2)
lines((model2$pred + 1.96*model2$se), col="red", lty=2, lwd=2)
lines((model2$pred - 1.96*model2$se), col="red", lty=2, lwd=2)
```



check residual assumption: iid, normality
```{r}
plot(as.vector(time(Y)), auto$residual, xlab="Days")
abline(h=0)
```
residual assumption: independence \ 

H0: The data are independently distributed 
Ha: The data are not independently distributed; they exhibit serial correlation.
```{r}
# Box-Pierce test and Ljung-Box test
Box.test(auto$residuals, lag=40, fitdf=2)
Box.test(auto$residuals, lag=40, fitdf=1)
Box.test(auto$residuals, lag=40, fitdf=0)
```

```{r}
acf(auto$residuals, lag.max=40)
rho.hat <- acf(auto$residuals, lag.max=40, plot=F)$acf[-1]
n <- length(auto$residuals)
h <- length(rho.hat)
q <- n*(n+2)*sum(rho.hat^2/(n-(1:h)))
q
pchisq(q, df=h, lower=F)

```

Check normality
```{r}
qqnorm(auto$residuals)
hist(auto$residuals)
```









```{r}
auto2 <- auto.arima(Y, ic="aicc", trace=T)
auto2
```

























## 4.4 What if we fit the model directly from OT

```{r}
#library(forecast)
OT.fit1 <- auto.arima(OT, trace=T, ic="aicc")
OT.fit1
```





```{r}
num <- length(OT)
# AIC
AIC(OT.fit1)/num - log(2*pi)
# BIC
AIC(OT.fit1, k=log(num))/num - log(2*pi)
# AICc
log(sum(resid(OT.fit1)^2)/num) + (num +5)/(num -5 -2)
```




```{r}
model2 <- predict(OT.fit1, n.ahead=10)
model2
```

```{r}
n <- length(Y)
L <- min(Y)
U <- max(Y)
plot(Y, xlim=c(1, n+3), ylim=c(L,U), type="line")
abline(h=0)

lines(model2$pred, col="red", lwd=2)
lines((model2$pred + 1.96*model2$se), col="red", lty=2, lwd=2)
lines((model2$pred - 1.96*model2$se), col="red", lty=2, lwd=2)
```



check residual assumption: iid, normality
```{r}
plot(as.vector(time(OT)), OT.fit1$residual, xlab="Days")
abline(h=0)
```
residual assumption: independence \ 

H0: The data are independently distributed 
Ha: The data are not independently distributed; they exhibit serial correlation.
```{r}
# Box-Pierce test and Ljung-Box test
Box.test(OT.fit1$residuals, lag=40, fitdf=2)
Box.test(OT.fit1$residuals, lag=40, fitdf=1)
Box.test(OT.fit1$residuals, lag=40, fitdf=0)
```

```{r}
acf(OT.fit1$residuals, lag.max=40)
rho.hat <- acf(OT.fit1$residuals, lag.max=40, plot=F)$acf[-1]
n <- length(OT.fit1$residuals)
h <- length(rho.hat)
q <- n*(n+2)*sum(rho.hat^2/(n-(1:h)))
q
pchisq(q, df=h, lower=F)

```

Check normality
```{r}
qqnorm(OT.fit1$residuals)
hist(OT.fit1$residuals)
```




































```{r}
OT.fit2 <- arima(OT, order=c(8,0,0), include.mean=F, method="ML")
OT.fit2
```

```{r}
OT.fit2 <- arima(OT, order=c(0,0,2), include.mean=F, method="ML")
OT.fit2
```




# 5 ARFIMA

```{r}
acf(OT, lag.max=100)
```


Long Memory process\









































