---
title: "STAT672_ Time Series Project"
output:
  html_document:
    df_print: paged
---

# Load Libraries
```{r}
library(usethis)
library(devtools)
library(ggplot2)
library(ggfortify)
library(dplyr)
library(magrittr)
library(lubridate)
library(forecast)
library(tidyverse)
library(mt)
library(lmtest)
library(outliers)
library(EnvStats)

```

# Load Data for station 1 

```{r}
getwd()
```

```{r}
oiltemp_1 <- read.csv("./oil station 1 hourly average for each day.csv",header=TRUE)
head(oiltemp_1, 10)
```

```{r}
summary(oiltemp_1)
```

```{r}
colnames(oiltemp_1)
```

```{r}
Feature_X<-as.matrix(cbind(oiltemp_1$HUFL,oiltemp_1$HULL, oiltemp_1$MUFL, oiltemp_1$MULL, oiltemp_1$LUFL, oiltemp_1$LULL, oiltemp_1$OT))
par(pairs(Feature_X))
```



```{r}
qqPlot(oiltemp_1$OT)
```


```{r}
ggplot(oiltemp_1, aes(x=TimeID)) + 
  geom_line(aes(y=OT), color="darkblue")
```


```{r}
xt1 <- oiltemp_1$OT
plot(xt1)
abline(h=0)
```

```{r}
#abline(h=0)
acf(xt1, lag.max=40)
```

```{r}
qqnorm(xt1)
```


```{r}
hist(xt1)
```

# Classical Regression 

```{r}
fit1 <- lm(xt1 ~ time(xt1))
summary(fit1)
```

```{r}
anova(fit1)
```



```{r}
plot(xt1, type="o", ylab="Oil Temperature 1")
abline(fit1)
```

```{r}
plot(as.vector(time(xt1)), fit1$res, xlab="Days")
abline(h=0)
```

oiltemp_1$HUFL + oiltemp_1$HULL + oiltemp_1$MUFL + oiltemp_1$MULL + oiltemp_1$LUFL + oiltemp_1$LULL
```{r}
fit2 <- lm(xt1 ~ oiltemp_1$HUFL + oiltemp_1$HULL + oiltemp_1$MUFL + oiltemp_1$MULL + oiltemp_1$LUFL + oiltemp_1$LULL + time(xt1))
summary(fit2)
```

```{r}
anova(fit2)
```

```{r}
plot(xt1, type="o", ylab="Oil Temperature 1")
abline(fit2)
```


```{r}
fit3 <- lm(xt1 ~ oiltemp_1$MUFL + oiltemp_1$LUFL + time(xt1))
summary(fit3)
```

```{r}
plot(xt1, type="o", ylab="Oil Temperature 1")
abline(fit3)
```

# Log Transformation 
```{r}
plot(log(xt1), main="log(oil temperature 1)", ylab="")

```


```{r}
hist(xt1)
```

```{r}
hist(log(xt1))
```

# Differencing to remove trend and seasonality (note7)
```{r}
plot(xt1)
```

```{r}
diff_xt1 <- diff(xt1)
plot(diff_xt1, type="b", lwd=2, col="blue")
```

```{r}
acf(diff_xt1, lag.max = 40)
```

```{r}
pacf(diff_xt1, lag.max = 40)
```



# Building Models 

## AR(p)
by pacf, cuts off after lag(11)

```{r}
model.fit1 <- arima(diff_xt1, order=c(11, 0, 0))
model.fit1
```

```{r}
forecast(model.fit1, h=6)
plot(forecast(model.fit1, h=6))

```


## MA(q) 
by acf, cuts off after lag(2) 

```{r}
model.fit2 <- arima(diff_xt1, order=c(0, 0, 2))
model.fit2
```

```{r}
forecast(model.fit2, h=6)
plot(forecast(model.fit2, h=6))
```

## Search for p, q 
sample ACF and PACF plots are of little use when both p>0 and q>0 in ARMA model. \
we need to find a systematic way to choose p and q. Assuming white noise normally distributed, we choose the model which has the smallest AICc value. \ 

```{r}
n <- length(diff_xt1)
for (p in 0:5){
  for (q in 0:5){
    model.fit <- arima(diff_xt1, order = c(p,0,q), method="CSS-ML")
    aicc <- model.fit$aic - 2*(p+q+1) + 2*(p+q+1)*n/(n-p-q-2)
    cat("p= ", p, ", q= ", q, ", AICC= ", aicc, "code= ", model.fit$code, "\n")
    
  }
}
```


## auto.arima

```{r}
auto.arima(diff_xt1)
auto.arima(diff_xt1, d=0, ic="aicc", trace=T)
```


```{r}
auto.arima(diff_xt1, ic="aicc", trace=T)
```














































